---
title: "Star Trek Voice Commands"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(skimr)
library(here)
tuesdata <- tidytuesdayR::tt_load(2021, week = 34)
computer <- tuesdata$computer
```

```{r cleaning, eval=FALSE}
# load package to extract json file
library(jsonlite)

# set location of file
url <- "http://www.speechinteraction.org/TNG/teaearlgreyhotdataset.json"

# read json string in: can use parse_json(url(url)) - unsure about difference?
raw_json <- read_json(url)

# view list structure of json string
raw_json %>% 
  listviewer::jsonedit()

clean_df <- raw_json %>% 
  # multi-level list -> data-frame with list columns
  enframe() %>% 
  # use unnest functions from tidyr to convert all lists into a dataframe
  unnest_longer(value) %>% 
  unnest_wider(value) %>% 
  unnest_longer(type) %>% 
  unnest_longer(domain) %>% 
  unnest_longer(`sub-domain`) %>% 
  janitor::clean_names()

clean_df %>% write_csv("2021/2021-08-17/computer.csv")
```

```{r}
skim(computer)
```

```{r explore_char}
# tabulate
computer %>% 
  group_by(char) %>% 
  count %>% 
  arrange(desc(n))

computer$char %>% n_distinct()
computer$char %>% unique
```

There are many characters but on inspection some appear to be duplicates e.g. due to different syntax or spelling. Use stringr to manipulate string data to tidy this up.

```{r tidy_char}
## detect any character name containing "com" or "Com"
## create object containing variations of computer
com_variation <- 
  computer[str_detect(computer$char, "(com|Com)+"),] %>%
  distinct(char) %>% 
  .$char

# turn computer$char into a factor 
computer$char <- as_factor(computer$char)

# collapse all factor levels containing "com|Com" into single factor level of "Computer"
computer$char <- 
  fct_collapse(.f = computer$char, Computer = com_variation)

# remove extra information in parentheses (e.g V.O. or O.S)
## identify pattern of " (" 
## remove this pattern and any characters that follow it
## sort out a few factors where no pattern
computer$char <- 
  computer$char %>% 
  str_replace_all(pattern = "\\s+\\(+.+", replacement = "") %>% 
  as_factor %>% 
  fct_recode(Picard = "Jean-Luc",
             Troi = "Mrs. Troi",
             Riker = "Riker'S Voice")

#check levels
computer$char %>% levels
```

```{r visualise_char}
computer %>%
  # main characters only (top 10)
  mutate(char = fct_lump(computer$char, 10)) %>% 
  ggplot(aes(y = fct_infreq(char)))+
  geom_bar()
```

Indicates commonest characters in data set - maybe should split into people and computer. There is a variable {char_type} that can be used for this.

Continue to tidy data set then do this at the end and re-visualise.

```{r explore_type}
computer$type %>% unique
# duplicates with different cases used
# switch all to title case and change to factor
computer$type <- 
  computer$type %>% 
  str_to_title() %>% 
  as_factor()

computer %>% 
  ggplot()+
  geom_bar(aes(x = fct_rev(fct_infreq(type))))+
  coord_flip()
```

13 possible interactions. Divide into computer and individual comments

```{r breakdown_types}
computer %>% 
  ggplot()+
  geom_bar(aes(x = fct_rev(fct_infreq(type))))+
  coord_flip()+
  facet_wrap(~ char_type)
```

Different types of interaction are used by 'person' and 'computer' except *conversation* which occurs in both groups.

Explore pri-type

```{r compare_pritype_type}
computer$pri_type <- 
  computer$pri_type %>% 
  as_factor()

# compare factor levels
a <- sort(levels(computer$type))
b <- sort(levels(computer$pri_type))
identical(a,b)
```

No manipulation required to 'pri-type' variable

```{r visualise_pri_type}
computer %>% 
  ggplot()+
  geom_bar(aes(x = fct_rev(fct_infreq(pri_type))))+
  coord_flip()+
  facet_wrap(~ char_type)
```

```{r explore_domain}
# standardise syntax to avoid duplication
computer$domain %>%  %>% unique

computer$sub_domain %>% 
  as_factor() %>% 
  fct_recode(Holodeck = "Holodeck?") %>% 
  levels
```

```{r tidy_domain_subdomain}
computer$domain <- 
  computer$domain %>% str_to_title() %>% 
  as_factor()

computer$sub_domain <- computer$sub_domain %>% 
  as_factor() %>% 
  fct_recode(Holodeck = "Holodeck?")
```

```{r}
# explore how subdomains relate to domains

computer %>% 
  group_by(domain, sub_domain) %>% 
  count

# visualise
computer %>% 
  ggplot() +
  geom_bar(aes(x = fct_infreq(sub_domain))) +
  coord_flip()+
  facet_wrap(~ domain)

# individual plots may be clearer. will need a function and use map() 

```

```{r}
# one final group should be factor {char_type}
computer$char_type <- 
  as_factor(computer$char_type)
```

Split data into 2 sets - computer and person

```{r}
person_type <- computer %>%
  filter(char_type == "Person")

comp_type <- computer %>% 
  filter(char_type == "Computer") 
```
