---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(skimr)
```

## Import

```{r import dataset}
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')
```

Parsing failures: explore with **problems()** and **spec()**

```{r}
problems(lemurs)
spec(lemurs)
```

### Parsing failure

variable: **age_of_living_y = col_logical()**

occurs in rows 1324:13450

```{r}
mean(is.na(lemurs$age_of_living_y))
```

all imported as NA - need to re import using read_csv. I suspect first 1324 values contain a lot of NA values and **read_csv** has interpreted variable as logical rather than as double.

Options: Increase **max_guess**, or manually select variable type on import

**col_types = cols(age_of_living_y = col_double())**

```{r read_again_maxguess}
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv', guess_max = 2000)
```

```{r read_again_coltypes}
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv', 
                          col_types = cols(
                            age_of_living_y = col_double()))
```

```{r}
problems(lemurs)
spec(lemurs)
mean(is.na(lemurs$age_of_living_y))
```

## Tidying

```{r skimr}
skim(lemurs)
```

## Exploration

```{r taxon}
lemurs %>% 
  ggplot(aes(x = fct_infreq(taxon)))+
  geom_bar()+
  coord_flip()

```

### Top 5 Species of lemur

Select top 5 species and compare distributions of various variables

```{r top_five}
top_five <- lemurs %>% 
  group_by(taxon) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(5) %>% 
  pull(taxon)

top_five <- lemurs %>% 
  filter(taxon %in% top_five)
```

```{r sex_diff}
top_five %>% 
  ggplot(aes(x = taxon, fill = sex)) +
  geom_bar(position = "fill")
```

```{r birth_type}
top_five %>% 
  ggplot(aes(x = taxon, fill = birth_type)) +
  geom_bar()

lemurs %>% 
  ggplot(aes(x = fct_infreq(taxon), fill = birth_type)) +
  geom_bar() +
  coord_flip()
```

```{r current_res}
lemurs %>% 
  ggplot(aes(x = fct_infreq(taxon), fill = current_resident)) +
  geom_bar() +
  coord_flip()
```

### Litter Size

```{r litter_size}

lemurs %>% 
  mutate(taxon = fct_infreq(taxon)) %>% 
  ggplot()+
  geom_jitter(aes(x = taxon,
                   y = litter_size), alpha = 0.05)+
  coord_flip()
```

Populations with more individuals seem to have individuals that come from larger litter.

Is there a relationship between litter size and number of individuals

```{r}
lemurs %>% 
  group_by(taxon) %>% 
  summarise(mean_litter = mean(litter_size, na.rm = TRUE),
            population = n()) %>% 
  ggplot() +
  geom_point(aes(y = mean_litter, x = population))+
  geom_smooth(aes(y = mean_litter, x = population), method = "loess")
```

```{r}
lemurs %>% 
  ggplot()+
  geom_bar(aes(litter_size))

```

### Weight

Investigate weight at different ages for different species then compare

```{r}
top_five %>% 
  ggplot()+
  geom_point(aes(x = age_at_wt_mo, y = weight_g, colour = taxon), alpha = 0.1)
```

```{r}
top_five %>% 
  ggplot()+
  geom_point(aes(x = age_at_wt_y, y = weight_g), alpha = 0.1, size = 0.5)+
  facet_wrap(~ taxon)
```

```{r}
top_five %>% 
  filter(taxon == "PCOQ" & dlc_id == 1591) %>% 
  ggplot()+
  geom_point(aes(x = age_at_wt_d, y = weight_g))

  
```

Sabrina is an outlier - wt gain ++++ then declined rapidly in old age

```{r}
top_five %>% 
  filter(dlc_id == 6117) %>% 
  ggplot()+
  geom_point(aes(x = age_at_wt_y, y = weight_g))+
  geom_smooth(aes(x = age_at_wt_y, y = weight_g))
```

```{r}
top_five %>% 
  filter(dlc_id != 6117) %>% 
  ggplot()+
  geom_point(aes(x = age_at_wt_mo, y = weight_g, colour = age_category), alpha = 0.2)
```

### Life Expectancy

```{r select_life_exp_vars}
life_expectancy <- lemurs %>% 
  select(taxon, dlc_id, sex, dob, estimated_dob,birth_type, birth_institution,
         dod, age_at_death_y, age_of_living_y, age_last_verified_y,
         age_max_live_or_dead_y)
```

```{r test_MMUR}
mmur <- life_expectancy %>% 
  filter(taxon == "MMUR") %>% 
  unique()

view(mmur)
```

```{r}
mmur %>% skim
```

filter to include only those with known dob and known dod and those born in captivity.

```{r}
mmur <- mmur %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" & is.na(estimated_dob)) %>%
  select(-c(estimated_dob, birth_type, birth_institution))
```

```{r visualise_mmur}
mmur %>% 
  ggplot()+
  geom_histogram(aes(x = age_at_death_y, fill = sex), binwidth = 1)

```

```{r test_PCOQ}
pcoq <- life_expectancy %>% 
  filter(taxon == "PCOQ") %>% 
  unique() %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" & is.na(estimated_dob)) %>%
  select(-c(estimated_dob, birth_type, birth_institution))

pcoq %>% 
  ggplot()+
  geom_histogram(aes(x = age_at_death_y), binwidth = 2)

```

```{r all_lemurs_life_exp}
all_lemur_life_exp <- life_expectancy %>% 
  unique() %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" & is.na(estimated_dob)) %>%
  select(-c(estimated_dob, birth_type, birth_institution))

all_lemur_life_exp %>% 
  ggplot()+
  geom_density(aes(x = age_at_death_y))
```

```{r top_five_life_exp}
top_five_life_exp <- top_five %>%  
  select(taxon, dlc_id, sex, dob, estimated_dob,birth_type, birth_institution,
         dod, age_at_death_y, age_of_living_y, age_last_verified_y,
         age_max_live_or_dead_y) %>%
  unique() %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" & is.na(estimated_dob)) %>%
  select(-c(estimated_dob, birth_type, birth_institution))

top_five_life_exp %>%
  ggplot()+
  geom_density(aes(x = age_at_death_y, fill = taxon)) +
  facet_wrap(~ taxon)

```

```{r}
top_five_life_exp %>% 
  ggplot()+
  geom_density(aes(x = age_at_death_y), fill = "steel blue")
```

```{r}
top_five_life_exp %>% 
  ggplot()+
  geom_violin(aes(x = taxon, y = age_at_death_y, fill = taxon))
```

```{r}
top_five_life_exp %>% 
  mutate(taxon = fct_reorder(taxon , age_at_death_y, na.rm = TRUE)) %>% 
  ggplot()+
  geom_boxplot(aes(x = age_at_death_y, fill = taxon))
```
