---
title: "Formula 1 - Import and Tidying"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```

## Circuits

```{r}
circuits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/circuits.csv')

```

No parsing errors.

Explore distribution of circuits by country

```{r}

circuits$country %>% n_distinct()

circuits %>% 
  group_by(country) %>%
  ggplot()+ 
  geom_bar(aes(x = fct_rev(fct_infreq(country))))+
  coord_flip()+
  scale_y_continuous(breaks = seq(0, 11, 1))
```

34 different countries have F1 circuits that have held Grand Prix races since 1950. USA has the most individual circuits with 11.

I wonder how many Grand Prix have been held in each country?

Possibility of a geo-spatial plot showing number of races held at each circuit or in each country maybe. Could also do a scatter plot comparing number of circuits and number of grand prix.

## Constructors

```{r}
constructors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/constructors.csv')
```

No parsing errors.

Explore distribution of constructors by nationality.

```{r}
constructors$nationality %>% n_distinct()

constructors %>% 
  ggplot()+
  geom_bar(aes(x = fct_rev(fct_infreq(nationality))))+
  coord_flip()
```

The British dominate the constructors.

Maybe look at constructors from each country over time.

## Drivers

```{r}
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/drivers.csv')
```

No parsing errors.

Look at distribution of drivers

```{r}
drivers$nationality %>% n_distinct()

drivers %>% 
  ggplot()+
  geom_bar(aes(x = fct_rev(fct_infreq(nationality))))+
  coord_flip()
```

British and US drivers are most common followed by Italian, German, French, Brazilian and Argentinian. Change American-Italian and Argentinian_Italian to American and Argentinian respectively.

## Races

```{r}
races <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/races.csv', na = "\\N")
```

```{r}
skimr::skim(races)

unique(races$time)
typeof(races$time)
```

Time parsed as character vector - due to "//N" which I presume is a missing value. Checked with API interface and confirms this. Best to sort this out on reading into R. I have added to code and it now parses this vector correctly as 'difftime'.

```{r}
class(races$time)
```

```{r}
races %>% 
  group_by(year) %>% 
  count %>% 
  ggplot(aes(x = year, y = n))+
  geom_col(colour = "black", fill = "grey")
```

## Results

```{r}
results <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/results.csv', na = "\\N")
```

```{r}
skimr::skim(results)
```

Time parsed as character due to winners time + difference. \> 1 lap behind then no time recorded i.e. NA. Possibly correct this to give a time for all finishers within 1 lap of the winner.

Work on single race to start with: most recent

```{r warning=FALSE}
results1 <- results %>% 
  mutate(time_behind = hms::hms(as.numeric(str_replace(time, "\\+", ""))),
         time_behind = hms(time_behind)) %>% 
  mutate_at(vars(time), na_if, "^\\+") %>% 
  mutate(time = hms(.$time)) %>% 
  relocate(time_behind, .after = "time")
```

## Constructor results

```{r}
constructor_results <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/constructor_results.csv', na = "\\N")
```

```{r}
constructor_results$status %>% unique()
constructor_results %>% 
  filter(status == "D")
```

## Status

```{r}
status <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/status.csv')

```

Look at status per race over a season - differences between teams

Check distribution of status

```{r}
results %>% 
  group_by(statusId) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  left_join(status)
```

## Qualifying

```{r}
qualifying <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/qualifying.csv', na = "\\N")
```

## Pitstops

```{r}
pit_stops <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/pit_stops.csv', na = "\\N", col_types = cols(duration = col_time()))

pit_stops <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/pit_stops.csv', na = "\\N")

problems(pit_stops)
```

Duration data parsed as character. Mixed format - hms. Convert to duration with lubridate.

```{r}
pit_stops <- pit_stops %>% 
  #filter(duration < 60) %>% 
  mutate(duration = as.numeric(milliseconds(milliseconds)))


```

```{r}
skimr::skim(pit_stops)
```

Distribution of pit_stop duration

```{r}
pit_stops %>% 
  ggplot()+
  geom_histogram(aes(duration), binwidth = 0.05)+
  scale_x_log10()

pit_stops %>% 
  ggplot()+
  geom_boxplot(aes(duration))+
  scale_x_log10()
```

## Lap-times

```{r}
lap_times <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/lap_times.csv')
```

```{r}
lap_times <- lap_times %>% 
  mutate(seconds = milliseconds/1000)
```

```{r}
skimr::skim(lap_times)
```

## Driver Standings

```{r}
driver_standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/driver_standings.csv', col_types = cols(
  driverStandingsId = col_double(),
  raceId = col_double(),
  driverId = col_double(),
  points = col_double(),
  position = col_double(),
  positionText = col_character(),
  wins = col_double()
))
```

```{r}
driver_standings %>% 
  group_by(positionText) %>% 
  count()
```

What is position_text and position? Why up to 99 and what is D and -?

## Constructor standings

```{r}
constructor_standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/constructor_standings.csv')

```

```{r}
constructor_standings %>% 
  group_by(positionText) %>% 
  count() %>% arrange(n)
```

## Seasons

```{r}
seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/seasons.csv')
```

\
