---
title: "Lemurs"
output: html_notebook
---

## Set Up

```{r set_up}
library(tidyverse)
library(skimr)
```

## Import

```{r}
# use tidytuesdayR package to read 'taxonomy' data only 
## parsing failure with lemurs data
taxonomy <- tidytuesdayR::tt_load(2021, week = 35, download_files = "taxonomy") %>% .$taxonomy

# read lemurs file sorting out parsing failure
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv', 
                          col_types = cols(
                            age_of_living_y = col_double()))
```

## Tidying and Transformation

Join both datasets so can use common names rather than 'taxon' codes

```{r}
# join both datasets so can use common names rather than 'taxon' codes
lemurs <- lemurs %>% 
  left_join(taxonomy) %>% 
  select(common_name, everything())
```

Group_together data for each individual lemur

```{r}
lemurs_species <- distinct(lemurs, dlc_id, .keep_all = TRUE)
```

There are multiple species which will make comparisons difficult when visualising. Create new object - a data-frame containing data for the 6 commonest species.

Visualise distribution first

```{r}
lemurs_species %>% 
  ggplot(aes(x = fct_infreq(common_name)))+
  geom_bar()+
  coord_flip()
```

NA and hybrid are common 'species'

Explore why NA and what is 'hybrid'

How many species in data set vs species in taxonomy set

```{r}
# count number of different taxon codes in each data frame
lemurs_species %>% 
  distinct(taxon) %>% nrow()
taxonomy %>% 
  distinct(taxon) %>% nrow()

# check if NAs in both sets or just the one
sum(is.na(lemurs_species$common_name))
sum(is.na(taxonomy$common_name))

# need to explore taxonomy dataset and look for error ?taxon codes that don't match
anti_join(lemurs_species, taxonomy) %>% 
  distinct(taxon) %>% .$taxon
```

*taxonomy* set refers to: 'CMEAD' - Cheirogaleus medius - "Fat-tailed dwarf lemur"

However *lemurs* data-set uses *taxon* code 'CMED'

Need to recode this so that they match and then re_join the data-frames.

```{r}
taxonomy$taxon <- 
  taxonomy$taxon %>% 
  fct_recode(CMED = "CMEAD")

lemurs <- lemurs %>% 
  left_join(taxonomy) %>% 
  select(common_name, everything())
```

Explore 'hybrid'

```{r}

```

```{r}
top_five <- lemurs %>% 
  group_by(taxon) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(5) %>% 
  pull(taxon)

top_five <- lemurs %>% 
  filter(taxon %in% top_five) 
```

## Exploration

### Descriptive statistics

```{r}
skim(lemurs)
```

```{r}
descriptive <- lemurs %>% 
  select(c(birth_month, age_at_death_y, litter_size, n_known_offspring)) %>%
  psych::describe() %>% 
  select(-(c(vars)))

descriptive
```

## Visualisation

#### Distribution of species

```{r}
lemurs %>% 
  ggplot(aes(x = fct_infreq(common_name)))+
  geom_bar()+
  coord_flip()
```

#### Birth Month

Explore distribution of months of birth.

```{r}
lemurs %>% 
  mutate(birth_month = as.integer(birth_month)) %>% 
  ggplot()+
  geom_density(aes(x = birth_month))+
  scale_x_continuous(breaks = c(1:12), labels = c(1:12))
```

Birth month is a discrete variable so convert numbers to month abbreviation and re-plot as a bar chart. Will need to produce an ordered factor for birth month and exclude NA. Also filter for captive born only and exclude any estimated dob.

```{r}
lemurs %>% 
  drop_na(birth_month) %>% 
  filter(birth_type == "CB" & is.na(estimated_dob)) %>% 
  mutate(birth_month = month.abb[birth_month],
         birth_month = factor(birth_month, 
                              levels = month.abb, 
                              ordered = TRUE)) %>% 
  ggplot(aes(x = birth_month))+
  geom_bar()
```

#### Age at death

Filter for those with known dob and known dod

```{r}
lemurs %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" 
         & is.na(estimated_dob)) %>% 
  ggplot(aes(x = age_at_death_y))+
  geom_histogram(binwidth = 1, colour = "white", fill = "brown")+
  geom_vline(xintercept = 12.05)
```

Multi-modal distribution - is this due to different mortality rates at different ages or is there a difference between species? Can look at data for top 6 species and compare rates.

```{r}
top_five %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" 
         & is.na(estimated_dob)) %>% 
  filter(!is.na(common_name) & common_name != "hybrid") %>% 
  ggplot(aes(x = age_at_death_y))+
  geom_histogram(aes(fill = common_name), binwidth = 1)
```
