---
title: "Lemurs"
output: html_notebook
---

## Set Up

```{r set_up}
library(tidyverse)
library(skimr)
library(patchwork)
library(cowplot)
library(ggridges)
```

## Import

```{r}
# use tidytuesdayR package to read 'taxonomy' data only 
## parsing failure with lemurs data
taxonomy <- tidytuesdayR::tt_load(2021, week = 35, download_files = "taxonomy") %>% .$taxonomy

# read lemurs file sorting out parsing failure
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv', 
                          col_types = cols(
                            age_of_living_y = col_double()))
```

## Tidying and Transformation

### Join datasets

Join both datasets so can use 'common names' rather than 'taxon' codes. First attempt resulted in missing values so I explored and corrected the differences between the two sets of data and re_joined them.

```{r}
# need to explore taxonomy dataset and look for error ?taxon codes that don't match
anti_join(lemurs, taxonomy) %>% 
  distinct(taxon) %>% .$taxon

# taxon = "CMED" in one set "CMEAD" in the other --> NA when left join
# recode so these match
taxonomy$taxon <- 
  taxonomy$taxon %>% 
  fct_recode(CMED = "CMEAD")

# join both datasets so can use common names rather than 'taxon' codes
lemurs <- lemurs %>% 
  left_join(taxonomy) %>% 
  select(common_name, everything())
```

### Single row for each individual lemur

Grouped together data for each individual lemur so each lemur is represented by a single row. Multiple rows exist due to weight data for different ages. Can use original dataset to look at weight data and this new data_frame to look at individual lemurs.

There are 2 different hybrid species both coded as 'hybrid' so they are joined together when grouping by 'common' name. This distorts the data so I chose to exclude hybrid species from my analysis.

```{r}
lemurs_distinct <- distinct(lemurs, dlc_id, .keep_all = TRUE) %>% 
  filter(common_name != "hybrid")
```

### Choose top species

There are 27 different species which will make comparisons difficult when visualising.

Create new object - a data-frame containing data for the *n* commonest species.

```{r}
n <- 5

top_species <- lemurs_distinct %>%  
  group_by(common_name) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(n) %>% 
  pull(common_name)

top_species<- lemurs_distinct %>% 
  filter(common_name %in% top_species) 
```

## Exploration

### Descriptive statistics

```{r}
skim(lemurs)
```

```{r}
descriptive <- lemurs %>% 
  select(c(birth_month, age_at_death_y, litter_size, n_known_offspring)) %>%
  psych::describe() %>% 
  select(-(c(vars)))

descriptive
```

## Visualisation

### Distribution of species

#### Overall

```{r}
lemurs_distinct %>% 
  ggplot(aes(x = fct_infreq(common_name)))+
  geom_bar()+
  coord_flip()
```

#### Top n species

```{r}
top_species %>% 
  ggplot(aes(x = fct_rev(fct_infreq(common_name))))+
  geom_bar()+
  coord_flip()
```

#### Birth Month

Explore distribution of months of birth.

Birth month is a discrete variable so convert numbers (integers) to month abbreviation and plot as a bar chart. Will need to produce an ordered factor for birth month and exclude missing values. Also exclude any estimated dob.

```{r}
lemurs_distinct %>% 
  drop_na(birth_month) %>% 
  filter(is.na(estimated_dob)) %>% 
  mutate(birth_month = month.abb[birth_month],
         birth_month = factor(birth_month, 
                              levels = month.abb, 
                              ordered = TRUE)) %>% 
  ggplot(aes(x = birth_month))+
  geom_bar()
```

#### Age at death

Filter for those with known dob and known dod. This is to ensure accurate data. These are all captive born so data only represents captive born lemurs

```{r}
life_exp <- top_species %>% 
  filter(!is.na(dob) & !is.na(dod) & is.na(estimated_dob)) 

life_exp[life_exp$birth_type != "CB",]

life_exp %>% 
  ggplot(aes(x = age_at_death_y))+
  geom_histogram(binwidth = 1, colour = "white", fill = "brown")+
  geom_vline(xintercept = mean(life_exp$age_at_death_y, na.rm = TRUE))
```

Multi-modal distribution - is this due to different mortality rates at different ages or is there a difference between species?

```{r}
life_exp %>% 
  ggplot(aes(x = age_at_death_y))+
  geom_density(aes(fill = common_name))+
  facet_wrap(~ common_name, ncol = 1)
```

Produce a plot of descriptive statistics for each of commonest species of lemur

```{r}
life_exp %>% 
  filter(!is.na(dob) & !is.na(dod) & birth_type == "CB" 
         & is.na(estimated_dob)) %>%
  group_by(common_name) %>% 
  summarise(mean_life_expectancy = mean (age_at_death_y, na.rm = TRUE)) %>% 
  mutate (common_name = fct_reorder(common_name, mean_life_expectancy)) %>% 
  ggplot()+
  geom_col(aes(x = common_name, y = mean_life_expectancy))+
  coord_flip()
```

```{r}
life_exp %>%
  mutate(common_name = fct_reorder(common_name, age_at_death_y)) %>% 
  ggplot()+
  geom_boxplot(aes(fill = common_name, x = common_name, y = age_at_death_y))
```

## Final Plots

Add colour

Tidy up labels etc.

Group together using patchwork package

Annotate as needed

### Generic settings

```{r}
lemur_pal <- c("#ccb79d", "#6e5337", "#687277", "#949992", "#c69d60")

mean_age <- mean(life_exp$age_at_death_y) %>% round(2)

life_exp <- life_exp %>%
  mutate(common_name = fct_reorder(common_name, age_at_death_y))
```

### Individual plots

```{r histogram_overall}
a

a <- life_exp %>% 
  ggplot(aes(x = age_at_death_y))+
  geom_histogram(binwidth = 1, 
                 colour = "black", 
                 fill = "#ccb79d")+
  labs(x = NULL, y = NULL)+
  theme_minimal_grid()
  
a
  
geom_segment(aes(x = mean_age,
                 xend = mean_age,
                 y = 0,
                 yend = 75),
             linetype = "dashed")+
  
labs(title = "Life expectancy has a multimodal distribution",
     subtitle = "More lemurs die in first year of life than at any other age")+
scale_x_continuous(name = "Age at death (yrs)")+
  
```

```{r barchart}
b <- life_exp %>%
  group_by(common_name) %>% 
  summarise(mean_life_expectancy = mean(age_at_death_y, na.rm = TRUE)) %>% 
  ggplot()+
  geom_col(aes(x = common_name, y = mean_life_expectancy, fill = common_name), 
           colour = "black")+
  coord_flip()+
  scale_fill_manual(values = lemur_pal)+
  labs(x = NULL, y = NULL)+
  theme_minimal_grid()+
  theme(legend.position = "none")

b
```

```{r boxplot}
c <- life_exp %>%
  ggplot()+
  geom_boxplot(aes(x = common_name, y = age_at_death_y,
                   fill = common_name))+ 
  coord_flip()+
  scale_fill_manual(values = lemur_pal)+
  labs(x = NULL, y = NULL)+
  theme_minimal_grid()+
  theme(legend.position = "none")

c
```

```{r ridge_plot}

d <- life_exp %>% 
    ggplot()+
    geom_density_ridges(aes(x = age_at_death_y, 
                            y = common_name,
                            height = ..density..,
                            fill = common_name,
                            scale = 0.95),
                        stat = "density") +
  scale_fill_manual(values = lemur_pal)+
  scale_y_discrete(breaks = NULL,
                       labels = NULL)+
  labs(y = NULL)+
  theme_minimal_vgrid()+
  theme(legend.position = "none")

d

```
